Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de Rham complex over convex domains and domains with shellable triangulations

### 2024-10-01 main.tex
more updates...

    Now we discuss the second recursive estimate. 
    Suppose again that $1 \leq m \leq M$. 
    Recall that we write $U_{F_m} := T_m \cup T_{m-1}$. 
    We use the Poincar\'e--Friedrichs inequality over $U_{F_{m}}$,
    as given in Lemma~\ref{lemma:poincarefriedrichsoverfacepatch}, 
    to find $w_{F_m} \in W^{1,p}(U_{F_m})$ such that $\nabla w_{F_m} = \nabla u$ over $U_{F_m}$ and 
    \begin{align}\label{math:wFm_inequality}
        \| w_{F_m} \|_{L^{p}(U_{F_m})} \leq C_{{\rm{PF}},U_{F_m},p} \| \nabla w_{F_m} \|_{L^{p}(U_{F_m})}.
    \end{align}
%     \cye{Since $\nabla (w|_{U_{F_m}}) = \nabla u_{F_{m}} = \nabla (u|_{U_{F_m}})$, 
    We can define the constant
    \begin{align*}
    c_{m} := w|_{U_{F_m}} - w_{F_{m}}.
    \end{align*}
    We now observe that 
    \begin{gather*}
        \| w \|_{L^{p}(T_m)}
        \leq 
        \| w_{F_m} \|_{L^{p}(T_m)}
        +
        \| c_{m} \|_{L^{p}(T_m)},
        \\
        \| c_{m} \|_{L^{p}(T_m)}
        = 
        \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
        \| c_{m} \|_{L^{p}(T_{{m-1}})},
        \\ 
        \| c_{m} \|_{L^{p}(T_{{m-1}})}
        \leq 
        \| w \|_{L^{p}(T_{{m-1}})} + \| w_{F_{m}} \|_{L^{p}(T_{{m-1}})} 
        .
    \end{gather*}
    In combination, 
    \begin{align*}
        \| w \|_{L^{p}(T_m)}
        &
        \leq 
        \| w_{F_m} \|_{L^{p}(T_m)}
        \\&\qquad 
        +
        \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
        \| w_{F_{m}} \|_{L^{p}(T_{{m-1}})}
        +
        \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
        \| w \|_{L^{p}(T_{{m-1}})}
        .
    \end{align*}
    We sum the two integrals of $w_{F_m}$. 
    % 
    % 1/q = 1 - 1/p = (p-1)/p 
    When $1 < p < \infty$, recalling the complementary exponent $q = p/(p-1) \in (1,\infty)$, 
    we use H\"older's inequality to verify 
    \begin{align*}
        \| w \|_{L^{p}(T_m)}
        &\leq 
        \left( 1 + \frac{ \vol(T_m)^{\frac q p} }{ \vol(T_{{m-1}})^{\frac q p} } \right)^{\frac 1 q}
        \left( 
            \| w_{F_{m}} \|_{L^{p}(T_m)}^{p}
            +
            \| w_{F_{m}} \|_{L^{p}(T_{{m-1}})}^{p}
        \right)^{\frac 1 p}
        +
        \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
        \| w \|_{L^{p}(T_{{m-1}})}
        \\&\leq 
        \left( 1 + \frac{ \vol(T_m)^{\frac q p} }{ \vol(T_{{m-1}})^{\frac q p} } \right)^{\frac 1 q}
        \| w_{F_{m}} \|_{L^{p}(U_{F_m})} 
        +
        \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
        \| w \|_{L^{p}(T_{{m-1}})}
        .
    \end{align*}
    In the special case $p=1$,
    \begin{align*}
        \| w\|_{L^{1}(T_m)}
        \leq 
        \max\left(
            1, \frac{ \vol(T_m) }{ \vol(T_{{m-1}}) } 
        \right)
        \| w_{F_m} \|_{L^{1}(U_{F_m})}
        +
        \frac{ \vol(T_m) }{ \vol(T_{{m-1}}) }
        \| w \|_{L^{1}(T_{{m-1}})}
        .
    \end{align*}
    In the limit case $p=\infty$, 
    \begin{align*}
        \| w \|_{L^{\infty}(T_m)}
        \leq 
        2
        \| w_{F_m} \|_{L^{\infty}(U_{F_m})}
        +
        \| w \|_{L^{\infty}(T_{{m-1}})}
        .
    \end{align*}
    The local inequality~\eqref{math:wFm_inequality} now provides the second recursive estimate. 
    The proof is complete.
\end{proof}

@@

Suppose that $\calT$ is an $n$-dimensional simplicial complex and we have an enumeration of the $n$-simplices $T_{0}, T_{1}, T_{2}, \dots \in \subsimplex_{n}(\calT)$.
For any enumeration, we call 
\begin{align*}
    I_m 
    := 
    \big( 
        T_{0} \cup T_{1} \cup \dots \cup T_{m} 
    \big) 
    \cap 
    T_{m+1}
\end{align*}
the $m$-th \textit{interface set}. 
% Note that this intersection is automatically a subcomplex of the boundary complex of the last simplex $T_{m}$.
We call the enumeration a \emph{shelling} if each interface set $I_m$ is a triangulated manifold of dimension $n-1$.

@@

\begin{lemma}\label{lemma:shell_triang_man}
    Let $\calT$ be an $n$-dimensional simplicial complex 
    with a shelling $T_{0}, T_{1}, T_{2}, \dots, T_{M}$,
    such that each simplex of dimension $n-1$ is contained in at most two simplices. 
    % DONE: what is non-branching?
    Then
    $$
        X_{m} := T_{0} \cup T_{1} \cup \dots \cup T_{m}
    $$ 
    is a triangulated manifold with boundary for all $0 \leq m \leq M$.
    % 
    In particular, $X_{m}$ is a topological $n$-ball when $m < M$, 
    and 
    $X_{M}$ is either a topological $n$-ball or topological $n$-sphere. 
\end{lemma}
\begin{proof}  
    We prove this claim by induction. 
    Certainly, if $\calT$ contains only one single $n$-simplex, then it is a shellable triangulation of a topological $n$-ball. 
    Suppose that $X_m := T_{0} \cup T_{1} \cup \dots \cup T_{m}$ is a topological $n$-ball and that $T_{m+1}$ is the next $n$-simplex in the shelling.
    By definition, $I_{m} := X_{m} \cap T_{m+1}$ is a submanifold of $\partial T_{m+1}$,
    and it is triangulated by some faces of $T_{m+1}$ and their subsimplices.

    Let $F$ be such a face. 
    By assumption, $F$ must be contained in exactly one $n$-simplex of $T_{0}, T_{1}, \dots, T_{m}$,
    and $F$ is in the boundary of $X_{m}$. We conclude that $I_{m}$ triangulates a submanifold of the boundary of $X_{m}$.

    On the one hand, 
    if $I_{m}$ is the entire boundary of $T_{m+1}$, 
    then it must also be the boundary of the topological $n$-ball $X_{m}$. 
    Hence $X_{m} \cup T_{m+1}$ must be a topological $n$-sphere and thus a manifold without boundary, which requires $m = M$.
    On the other hand, 
    if $I_{m}$ is a proper subset of the entire boundary of $T_{m+1}$, 
    then $X_{m} \cup T_{m+1}$ is still a topological $n$-ball.
\end{proof}

@@

\begin{lemma}\label{lemma:existenceofstar}
    Suppose that an $n$-dimensional manifold triangulation $\calT$ has a shelling $T_{0}, T_{1}, T_{2}, \dots, T_{M}$.
    For $0 \leq m < M$, write 
    $$
        X_{m} := T_{0} \cup T_{1} \cup \dots \cup T_{m},
        \qquad 
        I_{m} := X_{m} \cap T_{m+1}.
    $$ 
    Then $I_{m}$ is a union of $k$ different faces of $T$, $1 \leq k \leq n$, and the intersection of those faces is an interior simplex $S_{m} \in \calT$ of dimension $n-k$ that satisfies 
    $$
        \patch_{X_{m+1}}(S_{m}) = \patch_{\calT}(S_{m}).
    $$
\end{lemma}
\begin{proof}
    We know $I_{m}$ is a triangulated submanifold of the boundary of $T$, and so it must be a collection of $k$ faces of $T$, $1 \leq k \leq n$ ($k = n + 1$ can happen for the last enumerated simplex $m = M$ when $\calT$ is an $n$-sphere). $I_{m}$ also constitutes a local patch (star) of $(n-1)$-dimensional simplices around some simplex $S_{m}$ of dimension $n-k$ in $I_{m}$.
    By definition, $S_{m}$ is a boundary simplex of $X_{m}$, and it is an interior simplex of $X_{m+1}$. 
    But then $S_{m}$ cannot be a subsimplex of any of the simplices $T_{m+1}, \dots, T_{M}$,
    which means that $\patch_{X_{m+1}}(S_{m}) = \patch_{\calT}(S_{m})$. 
\end{proof}

@@

    In addition to that, by Proposition~\ref{proposition:pullbackestimate}:
    \begin{align*}
        \| w \|_{L^{p}(T_{m+1})}
        &\leq 
        |\det(\Jacobian \varphi)|^{\frac 1 p} 
        \matnorm{ \Jacobian \varphi^{-1} }^{k-1}
        \| \tilde w \|_{L^{p}(\hat T)}
        \\&
        \leq 
        |\det(\Jacobian \varphi)|^{\frac 1 p} 
        \matnorm{ \Jacobian \varphi^{-1} }^{k-1}
        \| \tilde w \|_{L^{p}(\tilde U)}
        ,
        \\
        \| \tilde g \|_{L^{p}(\tilde U)}
        &\leq 
        2^{\frac n p}
        \| \hat g \|_{L^{p}(\hat T)}
        \leq 
        2^{\frac n p}
        |\det( \Jacobian \varphi^{-1} )|^{\frac 1 p} 
        \matnorm{\Jacobian \varphi}^{k}
        \cdot 
        \| \cartan u \|_{L^{p}(T)}
        .
    \end{align*}
    The second inequality follows.
\end{proof}

Whenever $\calT$ is an $n$-dimensional triangulation, we use the abbreviation
\begin{align*}
    C_{\PF,\Gamma,k,p}(\calT) 
    :=
    \max\limits_{ \substack{ T \in \calT \; \dim(T)=n \\ 0 \leq l \leq n } }
    C_{\PF,\Gamma,l,k,p}(T)   
\end{align*}

Suppose that an $n$-dimensional triangulation $\calT$ has a shelling $T_{0}, T_{1}, \dots, T_{M}$. 
For each $1 \leq m \leq M$, there exists a simplex $S_{m} \in \calT$ 
such that $T_{m}$ is the last $n$-simplex in the shelling that belongs to $T_{m} \in \patch_{\calT}(S_{m})$.
We then let $U_{m}$ be the closure $\patch_{\calT}(S_{m}) \setminus T_{m}$,
and we call it the complement of $T_{m}$ in the star $\patch_{\calT}(S_{m})$.

The main result of this manuscript is the following theorem.

\begin{theorem}\label{theorem:poincarefriedrichsestimate:exterior}
    Let $\Omega \subseteq \bbR^{n}$ be a domain with a shellable triangulation $\calT$.
    Let $T_0, T_1, \dots, T_m$ be a shelling.
    Then for any $u \in W^{p}\Alt^{k}(\Omega)$ 
    there exists $w \in W^{p}\Alt^{k}(\Omega)$ with $\cartan u = \cartan w$ 
    and such that for each $T_{m} \in \calT$ in the shelling
    we have the following recursive estimate: 
    \begin{align*}
        \| w \|_{L^{p}(T_{m})}
%         &
%         \leq  
%         C_{\PF,\Gamma,k,p}(\calT) 
%         \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
%         \Cfive{n}{k}(\calT)^{k+1} 
%         \diam(T)^{k} 
%         \| \cartan u \|_{L^{p}(U_{m-1})}
%         \\&\qquad
%         + 
%         C_{\PF,\Gamma,k,p}(\calT) 
%         \diam(T)^{k} 
%         \| \cartan u      \|_{L^{p}(T_{m})} 
%         \\&\qquad
%         + 
%         \Cdetsechs{n}{k}(\calT)^{-\frac 1 p} 
%         \Cfive{n}{k}(\calT)^{k} 
%         \| w \|_{L^{p}(U_{m-1})}
%         \\
        &
        \leq  
        C_{\PF,\Gamma,k,p}(\calT) 
        \diam(T)^{k} 
        \left( 
            \| \cartan u      \|_{L^{p}(T_{m})} 
            +
            \Cfive{n}{k}(\calT)^{k+1} 
            \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
            \| \cartan u \|_{L^{p}(U_{m-1})}
        \right)
        \\&\qquad
        + 
        \Cfive{n}{k}(\calT)^{k} 
        \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
        \| w \|_{L^{p}(U_{m-1})}
        .
    \end{align*}
    \mwl{The notation $U_m$ should be settled in the definition of shellings.}
\end{theorem}

\begin{proof}
    This proof is similar to the proof of Theorem~\ref{theorem:poincarefriedrichsestimate:grad}.
    By assumption, there exists a shelling $T_0, T_1, \dots, T_m$. 
    We write $\Omega_m := \cup_{j=0}^{m} T_j$, which is a triangulated $n$-dimensional submanifold with boundary.

    Let $u \in W^{p}\Alt^{k}(\Omega)$. 
    First, there exists $w_0 \in W^{p}\Alt^{k}(T_{0})$ satisfying $\cartan w_0 = \cartan u$ over $T_{0}$ together with 
    \begin{gather*}
        \| w_0 \|_{L^{p}(T_{0})} \leq C_{{\rm{PF}},k,T_{0},p} \| \cartan u \|_{L^{p}(T_{0})}.
    \end{gather*}
    %Suppose that claim is true when the domain has $m$ simplices of dimension $n$. 
    Suppose that $0 < m \leq M$ such that there exists $w_{m-1} \in W^{p}\Alt^{k}(\Omega_{m-1})$ 
    with $\cartan w_{m-1} = \cartan u$ over $\Omega_{m-1}$. 
    By assumption, $T_{m}$ and $\overline\Omega_{m-1}$ share the interface $\Gamma_{m}$, which is a collection of faces of $T_{m}$. 
    In accordance to Lemma~\ref{lemma:existenceofstar}, adding $T_{m}$ completes a star in $\calT$ around some simplex, and we let $U_{m-1} \subseteq \Omega_{m-1}$ be the complement of $T_{m}$ in that newly completed star.

    We define $w'_{m} :=  \Xi_{m}^{\ast} w_{m-1} \in W^{p}\Alt^{k}(T_{m})$,
    where $\Xi_{m} : T_{m} \rightarrow U_{m-1}$ is the reflection mapping
    as in Proposition~\ref{proposition:starreflection}. 
    By construction, $w'_{m} \in W^{p}\Alt^{k}(T_{m})$ with 
    \begin{align*}
        \trace_F w_{m-1} = \trace_F w'_{m}
        . 
    \end{align*}
    There exists $w''_{m} \in W^{p}\Alt^{k}(T_{m})$ satisfying 
    \begin{align*}
        \cartan w''_{m} = \cartan u - \cartan w'_{m}, 
        \quad 
        \trace_{\Gamma_{m}} w''_{m} = 0.
    \end{align*}
    Indeed, such $w''_{m}$ exists because 
    \begin{align*}
        \trace_{\Gamma_{m}}( \cartan u - \cartan w'_{m} ) 
        &= 
        \trace_{\Gamma_{m}} \cartan u - \trace_{\Gamma_{m}} \cartan w'_{m}
        \\&= 
        \cartan \trace_{\Gamma_{m}} u - \cartan \trace_{\Gamma_{m}} w'_{m}
        = 
        \cartan \trace_{\Gamma_{m}} u - \cartan \trace_{\Gamma_{m}} w_{m-1}
        = 
        0
        .
    \end{align*}
    Setting $w_{m} := w_{m-1}$ over $\Omega_{m-1}$ and $w_{m} := w'_{m} + w''_{m}$ over $T_{m}$, 
    we find 
    \begin{gather*}
        \cartan w_{m} = \cartan w'_{m} + \cartan w''_{m} = \cartan w'_{m} + \cartan u - \cartan w'_{m} = \cartan u,
        \\
        \trace_{\Gamma_{m}} w_{m} = \trace_{\Gamma_{m}} w'_{m} = \trace_{\Gamma_{m}} w_{m-1}.
    \end{gather*}
    That means that $w_{m} \in W^{p}\Alt^{k}(\Omega_{m})$ with $\cartan w_{m} = \cartan u$ over $\Omega_{m}$.

>>>>

    Now we discuss the second recursive estimate. 
    Suppose again that $1 \leq m \leq M$. 
    Recall that we write $U_{F_m} := T_m \cup T_{m-1}$. 
    We use the Poincar\'e--Friedrichs inequality over $U_{F_{m}}$,
    as given in Lemma~\ref{lemma:poincarefriedrichsoverfacepatch}, 
    to find $w_{F_m} \in W^{1,p}(U_{F_m})$ such that $\nabla w_{F_m} = \nabla u$ over $U_{F_m}$ and 
    \begin{align}\label{math:wFm_inequality}
        \| w_{F_m} \|_{L^{p}(U_{F_m})} \leq C_{{\rm{PF}},U_{F_m},p} \| \nabla w_{F_m} \|_{L^{p}(U_{F_m})}.
    \end{align}
%     \cye{Since $\nabla (w_{|U_{F_m}}) = \nabla u_{F_{m}} = \nabla (u_{|U_{F_m}})$, 
    We can define the constant
    \begin{align*}
    c_{m} := w_{|U_{F_m}} - w_{F_{m}}.
    \end{align*}
    We now observe that 
    \begin{gather*}
        \| w \|_{L^{p}(T_m)}
        \leq 
        \| w_{F_m} \|_{L^{p}(T_m)}
        +
        \| c_{m} \|_{L^{p}(T_m)},
        \\
        \| c_{m} \|_{L^{p}(T_m)}
        = 
        \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
        \| c_{m} \|_{L^{p}(T_{{m-1}})},
        \\ 
        \| c_{m} \|_{L^{p}(T_{{m-1}})}
        \leq 
        \| w \|_{L^{p}(T_{{m-1}})} + \| w_{F_{m}} \|_{L^{p}(T_{{m-1}})} 
        .
    \end{gather*}
    In combination, 
    \begin{align*}
        \| w \|_{L^{p}(T_m)}
        &
        \leq 
        \| w_{F_m} \|_{L^{p}(T_m)}
        \\&\qquad 
        +
        \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
        \| w_{F_{m}} \|_{L^{p}(T_{{m-1}})}
        +
        \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
        \| w \|_{L^{p}(T_{{m-1}})}
        .
    \end{align*}
    We sum the two integrals of $w_{F_m}$. 
    % 
    % 1/q = 1 - 1/p = (p-1)/p 
    When $1 < p < \infty$, recalling the complementary exponent $q = p/(p-1) \in (1,\infty)$, 
    we use H\"older's inequality to verify 
    \begin{align*}
        \| w \|_{L^{p}(T_m)}
        &\leq 
        \left( 1 + \frac{ \vol(T_m)^{\frac q p} }{ \vol(T_{{m-1}})^{\frac q p} } \right)^{\frac 1 q}
        \left( 
            \| w_{F_{m}} \|_{L^{p}(T_m)}^{p}
            +
            \| w_{F_{m}} \|_{L^{p}(T_{{m-1}})}^{p}
        \right)^{\frac 1 p}
        +
        \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
        \| w \|_{L^{p}(T_{{m-1}})}
        \\&\leq 
        \left( 1 + \frac{ \vol(T_m)^{\frac q p} }{ \vol(T_{{m-1}})^{\frac q p} } \right)^{\frac 1 q}
        \| w_{F_{m}} \|_{L^{p}(U_{F_m})} 
        +
        \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
        \| w \|_{L^{p}(T_{{m-1}})}
        .
    \end{align*}
    In the special case $p=1$,
    \begin{align*}
        \| w\|_{L^{1}(T_m)}
        \leq 
        \max\left(
            1, \frac{ \vol(T_m) }{ \vol(T_{{m-1}}) } 
        \right)
        \| w_{F_m} \|_{L^{1}(U_{F_m})}
        +
        \frac{ \vol(T_m) }{ \vol(T_{{m-1}}) }
        \| w \|_{L^{1}(T_{{m-1}})}
        .
    \end{align*}
    In the limit case $p=\infty$, 
    \begin{align*}
        \| w \|_{L^{\infty}(T_m)}
        \leq 
        2
        \| w_{F_m} \|_{L^{\infty}(U_{F_m})}
        +
        \| w \|_{L^{\infty}(T_{{m-1}})}
        .
    \end{align*}
    The local inequality~\eqref{math:wFm_inequality} now provides the second recursive estimate. 
    The proof is complete.
\end{proof}

@@

Suppose that $\calT$ is an $n$-dimensional simplicial complex and we have an enumeration of the $n$-simplices $T_{0}, T_{1}, T_{2}, \dots \in \subsimplex_{n}(\calT)$.
For any enumeration, we call 
\begin{align*}
    \Gamma_m 
    := 
    \big( 
        T_{0} \cup T_{1} \cup \dots \cup T_{m} 
    \big) 
    \cap 
    T_{m+1}
\end{align*}
the $m$-th \textit{interface set}. 
% Note that this intersection is automatically a subcomplex of the boundary complex of the last simplex $T_{m}$.
We call the enumeration a \emph{shelling} if each interface set $\Gamma_m$ is a triangulated manifold of dimension $n-1$.

@@

\begin{lemma}\label{lemma:shell_triang_man}
    Let $\calT$ be an $n$-dimensional simplicial complex 
    with a shelling $T_{0}, T_{1}, T_{2}, \dots, T_{M}$,
    such that each simplex of dimension $n-1$ is contained in at most two simplices. 
    % DONE: what is non-branching?
    Then
    $$
        X_{m} := T_{0} \cup T_{1} \cup \dots \cup T_{m}
    $$ 
    is a triangulated manifold with boundary for all $0 \leq m \leq M$.
    % 
    In particular, $X_{m}$ is a topological $n$-ball when $m < M$, 
    and 
    $X_{M}$ is either a topological $n$-ball or topological $n$-sphere. 
\end{lemma}
\begin{proof}  
    We prove this claim by induction. 
    Certainly, if $\calT$ contains only one single $n$-simplex, then it is a shellable triangulation of a topological $n$-ball. Let $1 \leq m \leq M$ and suppose that 
    \begin{align*}
        X_{m-1} := T_{0} \cup T_{1} \cup \dots \cup T_{m-1}
    \end{align*}
    is a topological $n$-ball. Let $T_{m}$ is the next $n$-simplex in the shelling.
    By definition, $\Gamma_{m} := X_{m-1} \cap T_{m}$ is a submanifold of $\partial T_{m}$,
    and it is triangulated by some faces of $T_{m}$ and their subsimplices.

    Let $F$ be such a face. 
    By assumption, $F$ must be contained in exactly one $n$-simplex of $T_{0}, T_{1}, \dots, T_{m-1}$,
    and $F$ is in the boundary of $X_{m-1}$. 
    We conclude that $\Gamma_{m}$ triangulates a submanifold of the boundary of $X_{m-1}$.

    On the one hand, 
    if $\Gamma_{m}$ is the entire boundary of $T_{m}$, 
    then it must also be the boundary of the topological $n$-ball $X_{m-1}$. 
    Hence $X_{m-1} \cup T_{m}$ must be a topological $n$-sphere and thus a manifold without boundary, 
    which requires $m = M$.
    On the other hand, 
    if $\Gamma_{m}$ is a proper subset of the entire boundary of $T_{m}$, 
    then $X_{m-1} \cup T_{m}$ is still a topological $n$-ball.
\end{proof}

@@

\begin{lemma}\label{lemma:existenceofstar}
    Suppose that an $n$-dimensional manifold triangulation $\calT$ has a shelling $T_{0}, T_{1}, T_{2}, \dots, T_{M}$.
    For $0 \leq m \leq M$, write 
    $$
        X_{m} := T_{0} \cup T_{1} \cup \dots \cup T_{m}.
    $$ 
    For $1 \leq m \leq M$, write 
    $$
        \Gamma_{m} := X_{m-1} \cap T_{m}.
    $$ 
    Then $\Gamma_{m}$ is a union of $\ell$ different faces of $T_{m}$, $1 \leq k \leq n+1$.
    If $m < M$, then the intersection of those faces is an interior simplex $S_{m} \in \calT$ of dimension $n-k$ that satisfies 
    $$
        \patch_{X_{m}}(S_{m}) = \patch_{\calT}(S_{m}).
    $$
\end{lemma}
\begin{proof}
    We know $\Gamma_{m}$ is a triangulated submanifold of the boundary of $T_{m}$, 
    and so it must be a collection of $k$ faces of $T$, $1 \leq k \leq n+1$.
    Note that $k = n + 1$ can only happen for the last enumerated simplex, $m = M$, if $\calT$ triangulates an $n$-sphere. 
    $\Gamma_{m}$ also constitutes a local patch (star) of $(n-1)$-dimensional simplices around some simplex $S_{m}$ of dimension $n-k$ in $\Gamma_{m}$.
    By definition, $S_{m}$ is a boundary simplex of $X_{m}$, 
    and it is an interior simplex of $X_{m+1}$. 
    But then $S_{m}$ cannot be a subsimplex of any of the simplices $T_{m+1}, \dots, T_{M}$,
    which means that $\patch_{X_{m+1}}(S_{m}) = \patch_{\calT}(S_{m})$. 
\end{proof}

@@

    In addition to that, by Proposition~\ref{proposition:pullbackestimate}:
    \begin{align*}
        \| w \|_{L^{p}(T)}
        &\leq 
        |\det(\Jacobian \varphi)|^{\frac 1 p} 
        \matnorm{ \Jacobian \varphi^{-1} }^{k-1}
        \| \tilde w \|_{L^{p}(\hat T)}
        \\&
        \leq 
        |\det(\Jacobian \varphi)|^{\frac 1 p} 
        \matnorm{ \Jacobian \varphi^{-1} }^{k-1}
        \| \tilde w \|_{L^{p}(\tilde U)}
        ,
        \\
        \| \tilde g \|_{L^{p}(\tilde U)}
        &\leq 
        2^{\frac n p}
        \| \hat g \|_{L^{p}(\hat T)}
        \leq 
        2^{\frac n p}
        |\det( \Jacobian \varphi^{-1} )|^{\frac 1 p} 
        \matnorm{\Jacobian \varphi}^{k}
        \cdot 
        \| \cartan u \|_{L^{p}(T)}
        .
    \end{align*}
    \todo{MV: the norms of the Jacobians have to be fixed.}
    The second inequality follows.
    \todo{MV: Please a few more details.}
\end{proof}

Whenever $\calT$ is an $n$-dimensional triangulation, we use the abbreviation
\begin{align*}
    C_{\PF,\Gamma,k,p}(\calT) 
    :=
    \max\limits_{ \substack{ T \in \calT \\ \dim(T)=n \\ 0 \leq l \leq n } }
    C_{\PF,\Gamma,l,k,p}(T)
    .
\end{align*}
We settle the following notation for our main result. 
Suppose that an $n$-dimensional triangulation $\calT$ has a shelling $T_{0}, T_{1}, \dots, T_{M}$. 
For each $1 \leq m \leq M$, the interface of the new simplex to the previous intermediate triangulation is 
\begin{align*}
    \Gamma_{m} := T_{m} \cap \left( T_{0} \cup \dots \cup T_{m-1} \right)
\end{align*}
According to Lemma~\ref{lemma:existenceofstar}, 
there exists a simplex $S_{m} \in \calT$ 
such that $T_{m}$ is the last $n$-simplex in the shelling that belongs to $T_{m} \in \patch_{\calT}(S_{m})$.
In particular, $\patch_{\calT}(S_{m})$ is already included in the union of the simplices up to $T_m$. 
We write 
\begin{align*}
    U_{m} := \patch_{\calT}(S_{m}) \setminus T_{m}.
\end{align*}
for the complement of $T_{m}$ in the star $\patch_{\calT}(S_{m})$.

The main result of this manuscript for shellable triangulations of domains, not necessarily convex or even star-shaped, is the following theorem.

\begin{theorem}\label{theorem:poincarefriedrichsestimate:exterior}
    Let $\Omega \subseteq \bbR^{n}$ be a domain with a shellable $n$-dimensional triangulation $\calT$.
    Let $T_0, T_1, \dots, T_M$ be a shelling.
    Then for any $u \in W^{p}\Alt^{k}(\Omega)$ 
    there exists $w \in W^{p}\Alt^{k}(\Omega)$ with $\cartan w = \cartan u$ 
    and such that for each $T_{m} \in \calT$, $1 \leq m \leq M$, in the shelling
    we have the following recursive estimate: 
    \begin{align*}
        \| w \|_{L^{p}(T_{m})}
        &
        \leq  
        C_{\PF,T_{m},\Gamma_{m},k,p}(\calT) 
        \diam(T)^{k} 
        \left( 
            \| \cartan u      \|_{L^{p}(T_{m})} 
            +
            \Cfive{n}{k}(\calT)^{k+1} 
            \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
            \| \cartan u \|_{L^{p}(U_{m-1})}
        \right)
        \\&\qquad
        + 
        \Cfive{n}{k}(\calT)^{k} 
        \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
        \| w \|_{L^{p}(U_{m-1})}
        .
    \end{align*}
    \mwl{The notation $U_m$ should be settled in the definition of shellings.}
\end{theorem}
\begin{proof}
    This proof is similar to the proof of Theorem~\ref{theorem:poincarefriedrichsestimate:grad}.
    By assumption, there exists a shelling $T_0, T_1, \dots, T_m$. 
    We write $X_m := \cup_{j=0}^{m} T_j$, which is a triangulated $n$-dimensional submanifold with boundary.

    Let $u \in W^{p}\Alt^{k}(\Omega)$. 
    First, there exists $w_0 \in W^{p}\Alt^{k}(T_{0})$ satisfying $\cartan w_0 = \cartan u$ over $T_{0}$ together with 
    \begin{gather*}
        \| w_0 \|_{L^{p}(T_{0})} \leq C_{{\rm{PF}},k,T_{0},p} \| \cartan u \|_{L^{p}(T_{0})}.
    \end{gather*}
    %Suppose that claim is true when the domain has $m$ simplices of dimension $n$. 
    Suppose that $0 < m \leq M$ such that there exists $w_{m-1} \in W^{p}\Alt^{k}(X_{m-1})$ 
    with $\cartan w_{m-1} = \cartan u$ over $X_{m-1}$. 
    By assumption, $T_{m}$ and $X_{m-1}$ share the interface $\Gamma_{m}$, which is a collection of faces of $T_{m}$. 
    In accordance to Lemma~\ref{lemma:existenceofstar}, adding $T_{m}$ completes a star in $\calT$ around some simplex, and we let $U_{m-1} \subseteq X_{m-1}$ be the complement of $T_{m}$ in that newly completed star.

    We define $w'_{m} :=  \Xi_{m}^{\ast} w_{m-1} \in W^{p}\Alt^{k}(T_{m})$,
    where $\Xi_{m} : T_{m} \rightarrow U_{m-1}$ is the reflection mapping
    as in Proposition~\ref{proposition:starreflection}. 
    By construction, $w'_{m} \in W^{p}\Alt^{k}(T_{m})$ with 
    \begin{align*}
        \trace_F w_{m-1} = \trace_F w'_{m}
        . 
    \end{align*}
    There exists $w''_{m} \in W^{p}\Alt^{k}(T_{m})$ satisfying 
    \begin{align*}
        \cartan w''_{m} = \cartan u - \cartan w'_{m}, 
        \quad 
        \trace_{\Gamma_{m}} w''_{m} = 0.
    \end{align*}
    Indeed, such $w''_{m}$ exists because 
    \begin{align*}
        \trace_{\Gamma_{m}}( \cartan u - \cartan w'_{m} ) 
        &= 
        \trace_{\Gamma_{m}} \cartan u - \trace_{\Gamma_{m}} \cartan w'_{m}
        \\&= 
        \cartan \trace_{\Gamma_{m}} u - \cartan \trace_{\Gamma_{m}} w'_{m}
        = 
        \cartan \trace_{\Gamma_{m}} u - \cartan \trace_{\Gamma_{m}} w_{m-1}
        = 
        0
        .
    \end{align*}
    Setting $w_{m} := w_{m-1}$ over $X_{m-1}$ and $w_{m} := w'_{m} + w''_{m}$ over $T_{m}$, 
    we find 
    \begin{gather*}
        \cartan w_{m} = \cartan w'_{m} + \cartan w''_{m} = \cartan w'_{m} + \cartan u - \cartan w'_{m} = \cartan u,
        \\
        \trace_{\Gamma_{m}} w_{m} = \trace_{\Gamma_{m}} w'_{m} = \trace_{\Gamma_{m}} w_{m-1}.
    \end{gather*}
    That means that $w_{m} \in W^{p}\Alt^{k}(X_{m})$ with $\cartan w_{m} = \cartan u$ over $X_{m}$.
